Specialisation \& the Latitude-niche-breadth Hypothesis

### 2015-10-16 talk/ABC_2015.tex
revamped talk following confab with DBS

  \begin{frame}{Why are there more species in the tropics?}

@@

    \textbf{Alyssa R. Cirtwill},\\ Daniel B. Stouffer \& Tamara N. Romanuk

@@

      {\color{white}{\Large Latitude-Niche Breadth Hypothesis}}

      Envionmental stability? 
      Productivity? 
      Speciation rates?

@@

      {\Large Latitude-Niche Breadth Hypothesis}

      {\color{purple}Envionmental stability? Productivity?}
      Speciation rates?

@@

  \begin{frame}{Latitude-Niche Breadth Hypothesis}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_1.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
  \end{frame}

  \begin{frame}{Latitude-Niche Breadth Hypothesis}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_2.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
  \end{frame}

  \begin{frame}{Latitude-Niche Breadth Hypothesis}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_3.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
  \end{frame}

  \begin{frame}{Alternatively ...}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_4.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Davies \emph{et. al.} 2007. Productivity alters the scale dependence of the diversity-invasibility relationship. \emph{Ecology.}}
  \end{frame}

  \begin{frame}{Alternatively ...}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_5.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Davies \emph{et. al.} 2007. Productivity alters the scale dependence of the diversity-invasibility relationship. \emph{Ecology.}}
  \end{frame}

\section*{Food Webs}
  \begin{frame}{Does niche breadth vary with latitude?}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

@@

  \begin{frame}{Does niche breadth vary with latitude?}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

@@

@@

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/LittleRockLake.eps}

    195 food webs \\
    0-78$^\circ$ (absolute latitude) \\
    {\color{white} Mean generality (\textbf{G})}

@@

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/LittleRockLake.eps}

    195 food webs \\
    0-78$^\circ$ (absolute latitude) \\
    Mean generality (\textbf{G})

@@

  \begin{frame}{Niche breadth varies with species richness}

@@

@@

    \vspace{.3cm}
      {\color{white}{\Large      
      $G \approx aS^b$}}
    \end{center}

@@

  \begin{frame}{Does latitude affect scaling?}

@@

      \vspace{.3cm}
    \end{center}
      \hspace{.74cm}
      {\Large
      $G \approx aS^{\color{purple}b_0+b_1Latitude}$}
  \end{frame}

  \begin{frame}{Does latitude affect scaling?}
    \begin{center}
      \includegraphics*[width=.654\textwidth]{Figures/results/Gen_dots_vs_S_fitline_observed.eps}

      \vspace{.3cm}
    \end{center}
      \hspace{.74cm}
      {\Large
      $G \approx aS^{\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$}

\section*{Results}
  \begin{frame}{Does latitude affect scaling?}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/Gen_vs_S_marginal_axis.eps}

      \vspace{.8cm}
      ${\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$

@@

      \vspace{.8cm}
      ${\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$

@@

      \vspace{1cm}
      {\color{white} No effect of latitude on scaling\\
      Similar niche widths at all latitude\\
      \vspace{.54cm}        
      Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics}

@@

      \vspace{1cm}
      No effect of latitude on scaling\\
      {\color{white}Similar niche widths at all latitude\\
      \vspace{.54cm}      
      Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics}

      \vspace{1cm}
      No effect of latitude on scaling\\
      Similar niche widths at all latitude\\
      \vspace{.54cm}
      {\color{white} Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics}

      \vspace{1cm}
      No effect of latitude on scaling\\
      Similar niche widths at all latitude\\
      \vspace{.54cm}      
      Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics

@@

      \vspace{1cm}
      {\color{white} Scaling increases with latitude\\
      Narrower niches in the tropics\\
      \vspace{.54cm}      
      Supports Vazquez \& Stevens 2004\\}
      {\color{white} Narrower niches in the tropics}

@@

      \vspace{1cm}
      Scaling increases with latitude\\
      {\color{white} Narrower niches in the tropics\\
      \vspace{.54cm}      
      Supports Vazquez \& Stevens 2004\\}
      {\color{white} narrower niches in the tropics}

      \vspace{1cm}
      Scaling increases with latitude\\
      Narrower niches in the tropics\\
      \vspace{.54cm}     
      {\color{white} Supports Vazquez \& Stevens 2004\\}
      {\color{white} narrower niches in the tropics}

      \vspace{1cm}
      Scaling increases with latitude\\
      Narrower niches in the tropics\\
      \vspace{.54cm}
      Supports Vazquez \& Stevens 2004\\
      {\color{white} narrower niches in the tropics}

@@

  \begin{frame}{What's different in freshwaters?}
    \begin{block}{They're highly seasonal}

@@

  \begin{frame}{What's different in freshwaters?}
    \begin{block}{They're highly seasonal - except in NZ}

@@

        \includegraphics[width=3in]{Figures/sawcut.eps}

  \begin{frame}{What's different in freshwaters?}
    \begin{block}{They're dominated by fish}

      \begin{columns}
      \column{.5in}
      \column{1.9in}
        \hfill
        \includegraphics[width=.8\columnwidth]{Figures/anteater.eps}
      \column{.2in}
      \column{1.9in}
        \includegraphics[width=.8\columnwidth]{Figures/trout.eps}
      \column{.5in}
      \end{columns}
    \end{block}

\section*{Conclusion and Acknowledgements}
  \begin{frame}{So why are there more species in the tropics?}
    \begin{columns}
    \column{.5in}
    \column{2.5in}
      \begin{block}{Estuarine, marine, \& terrestrial}
        \begin{itemize}
          \item Broader niche space\\ in the tropics
          \item Similar specialisation\\ everywhere
        \end{itemize}
      \end{block}

    \begin{block}{Lakes \& Streams}
      \begin{itemize}
        \item Narrower niches\\ in the tropics
        \item Generality due to seasonality?
      \end{itemize}
    \end{block}

      \column{1.5in}
        \centering
        \includegraphics[height=.75in]{Figures/version2.eps}

        \includegraphics[height=.75in]{Figures/version1.eps}
    \column{.5in}
    \end{columns}

@@

  \begin{frame}{Latitude affects scaling in freshwaters}

@@

    {\Large $G \approx aS^{\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$}

>>>>

  \begin{frame}{Testing the latitude-niche breadth hypothesis}

@@

    \textbf{Alyssa R. Cirtwill}\\ Daniel B. Stouffer \& Tamara N. Romanuk

@@

      Many groups are more diverse in the tropics\\
      \vspace{.5cm}
      {\color{white} Environmental stability? Productivity?\\}

@@

      Many groups are more diverse in the tropics\\
      \vspace{.5cm}
      {\color{white}{Ey}}How?{\color{white}{Ey}}\\

@@

  \begin{frame}{Testing the latitude-niche breadth hypothesis}

      Many groups are more diverse in the tropics\\
      \vspace{.5cm}
      Environmental stability? Productivity?\\

  \begin{frame}{Testing the latitude-niche breadth hypothesis}

    \begin{block}{Latitude-niche breadth version 1}
    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_1.eps}
    \end{center}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
    \end{block}
  \end{frame}

  \begin{frame}{Testing the latitude-niche breadth hypothesis}

    \begin{block}{Latitude-niche breadth version 1}
    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_2.eps}
    \end{center}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
    \end{block}
  \end{frame}

  \begin{frame}{Testing the latitude-niche breadth hypothesis}

    \begin{block}{Latitude-niche breadth version 1}
    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_3.eps}
    \end{center}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
    \end{block}
  \end{frame}

  \begin{frame}{Testing the latitude-niche breadth hypothesis}
    \begin{block}{Latitude-niche breadth version 2}
    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_4.eps}
    \end{center}

    \tiny{Davies \emph{et. al.} 2007. Productivity alters the scale dependence of the diversity-invasibility relationship. \emph{Ecology.}}
    \end{block}
  \end{frame}

  \begin{frame}{Testing the latitude-niche breadth hypothesis}
    \begin{block}{Latitude-niche breadth version 2}
    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_5.eps}
    \end{center}

    \tiny{Davies \emph{et. al.} 2007. Productivity alters the scale dependence of the diversity-invasibility relationship. \emph{Ecology.}}
    \end{block}
  \end{frame}

\section*{Food Webs}
  \begin{frame}{Testing the latitude-niche breadth hypothesis}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

@@

    \begin{center}
      \color{white}{Generality : mean number of prey per species}
    \end{center}
  \end{frame}

  \begin{frame}{Testing the latitude-niche breadth hypothesis}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

@@

    \begin{center}
      \color{white}{Generality : mean number of prey per species}
    \end{center}
  \end{frame}

  \begin{frame}{Testing the latitude-niche breadth hypothesis}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

      \vspace{.25cm}
      Lower generality\\in the tropics

      \vspace{.25cm}
      Equal generality\\at all latitudes

    \begin{center}
      \color{DarkGreen}{Generality : mean number of prey per species}
    \end{center}

@@

    \begin{block}{Dataset}
    \begin{center}
      \includegraphics*[width=.75\textwidth]{Figures/LittleRockLake.eps}

      \vspace{.4cm}
    195 food webs \\
    5-188 species \\

    \begin{block}{Dataset}
    \begin{center}
      \includegraphics*[width=.75\textwidth]{Figures/results/latitude_hist.eps}

    \end{center}
    \end{block}

@@

    \begin{block}{Dataset}
    \begin{center}
      \includegraphics*[width=.75\textwidth]{Figures/results/generality_hist.eps}

    \end{center}
    \end{block}

@@

  \begin{frame}{Niche breadth scales with species richness}

@@

  \begin{frame}{Niche breadth scales with species richness}

    \vspace{.3cm}
      {\Large      
      log($G$) $\approx b$  log($a$)  log($S$) }
    \end{center}
  \end{frame}

 \begin{frame}{Niche breadth scales with species richness}

@@

    \vspace{.3cm}
      {\Large      
      $G \approx aS^b$ \hspace{.5in} $S \approx Latitude$}
    \end{center}

@@

  \begin{frame}{Does latitude affect scaling of niche breadth?}

@@

    \vspace{.3cm}
      {\Large      
      $G \approx aS^{\color{DeepPurple}b}$}
    \end{center}
  \end{frame}

  \begin{frame}{Does latitude affect scaling of niche breadth?}
    \begin{center}
      \includegraphics*[width=.654\textwidth]{Figures/results/Gen_dots_vs_S_fitline_observed.eps}

      \vspace{.3cm}
      {\Large
      $G \approx aS^{\color{DeepPurple}b_0+b_1Latitude}$}
    \end{center}
  \end{frame}

  \begin{frame}{Does latitude affect scaling of niche breadth?}
    \begin{center}
      \includegraphics*[width=.654\textwidth]{Figures/results/Gen_vs_S_expectations.eps}

      \vspace{.3cm}
      {\Large
      $G \approx aS^{\color{DeepPurple}b_0+b_1Latitude}$}
    \end{center}
  \end{frame}

\section*{Results}
  \begin{frame}{Does latitude affect scaling of niche breadth?}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/Gen_vs_S_marginal_axis.eps}

      \vspace{.3cm}
      {\Large
      $G \approx aS^{\color{DeepPurple}b_0+b_1Latitude}$}

@@

      \vspace{.3cm}
      {\Large
      $G \approx aS^{\color{DeepPurple}b_0+b_1Latitude}$}

@@

    \includegraphics*[width=.75\textwidth]{Figures/results/no_effect_interpretations.eps}

@@

    \vspace{1.02cm}

    \includegraphics*[width=.4\textwidth]{Figures/version2.eps}

@@

    \includegraphics*[width=.75\textwidth]{Figures/results/effect_interpretations.eps}

@@

    \vspace{1.02cm}

    \includegraphics*[width=.255\textwidth]{Figures/version1.eps}

@@

  % \begin{frame}{How do lakes and streams differ}
  %   \begin{block}{from terrestrial food webs?}

  %     \begin{columns}
  %       \column{.5in}

  %       \column{1.9in}
  %         \begin{center}
  %         \includegraphics[width=.8\textwidth]{Figures/anteater.eps}
  %         \end{center}
  %       \column{.2in}
  %       \column{1.9in}
  %         \begin{center}
  %         \includegraphics[width=.8\textwidth]{Figures/trout.eps}
  %         \end{center}
  %       \column{.5in}
  %     \end{columns}

  %       \vspace{.5cm}

  %     \begin{columns}
  %       \column{.5in}

  %       \column{1.9in}

  %        \begin{center}

  %       Gape-limitation!

  %       \vspace{.25cm}

  %       {\tiny Liem 1990, \emph{Am. Zool}}\\
  %       {\tiny Shurin \emph{et. al.} 2007, \emph{Ecol. Lett.}}
  %       \end{center}

  %       \column{.2in}

  %       \column{1.9in}

  %         \begin{center}
  %         \includegraphics[width=.5\columnwidth]{Figures/version1.eps}

  %         \end{center}

  %       \column{.5in}

  %     \end{columns}
  %   \end{block}
  % \end{frame}

  \begin{frame}{How do lakes and streams differ}
    \begin{block}{from estuarine \& marine food webs?}

@@

\section*{Conclusion and Acknowledgements}
  \begin{frame}{Evidence for the latitude-niche breadth hypothesis}
    \begin{block}{Estuarine, marine, \& terrestrial webs}
      \begin{columns}
      \column{2.5in}
        \begin{itemize}
          \item Broader niche space\\ in the tropics
          \item Similar specialisation\\ everywhere
        \end{itemize}
      \column{1in}

@@

        \includegraphics[height=.75in]{Figures/version2.eps}

      \column{.5in}
      \end{columns}
    \end{block}

    \begin{block}{Lake \& stream webs}
      \begin{columns}
      \column{2.5in}
      \begin{itemize}
        \item Narrower niches\\ in the tropics
        \item Specialisation\\ due to stability?
      \end{itemize}
      \column{1in}

        \includegraphics[height=.75in]{Figures/version1.eps}

      \column{.5in}
    \end{columns}
    \end{block}
  \end{frame}

@@

  \begin{frame}{Evidence for the latitude-niche breadth hypothesis}

@@

    {\Large $G \approx aS^{\color{DeepPurple}b_0+b_1Latitude}$}
